### 2018-05-04 markdown_whiteboard.Rmd
updated and cleaned data prep

---
output:
  pdf_document: default
---
We now have a trove of new data: data describing interactions between species, interactions that vary with the state of ecosystem. In general, the multivariate models perform well (i.e. achieve good preditive power).

```{r multivariate model stats,message=F,warning=F}
library(knitr)
library(tidyverse)
require(kableExtra)

x<-data.frame(spp=c(1,2,3),theta=c(3,4,5),rho=c(.89,.99,.32),mae=c(.4,.5,.6),
                lngstr=rep("this is a long string. this is a long string. this is a long string. this is a long string.this is a long string. this is a long string.this is a long string. this is a long string.this is a long string. this is a long string.this is a long string. this is a long string.",3))

knitr::kable(x,format = "latex",col.names=c("Species","Nonlinearity", "$\\rho$","MAE","String"),
             caption="Multivariate S-map Model Statistics, with nonlinearity,
             correlation of predictions with observations, and mean absolute error.",
             escape=F) %>%
  kable_styling(full_width = F) %>%
  column_spec(3,bold=T)%>%
  column_spec(5, width = "30em")

```